---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mizerUR

<!-- badges: start -->
<!-- badges: end -->

This is an extension package for the mizer package
https://sizespectrum.org/mizer/. Mizer has a size-structured planktonic
resource. With this extension you can also model any number of unstructured
resource components. Such unstructured components are appropriate whenever the
predation on these components is not size based.

Possible applications include the modelling of detritus as a resource for
detritivores, carrion as a resource for scavengers, or macroflora on which fish
can graze.

## Installation

You can install mizerUR from [GitHub](https://github.com/gustavdelius/mizerUR)
with:

``` r
devtools::install_github("gustavdelius/mizerUR")
```

## Example

The mizerUR package facilitates adding unstructured resource components to an
existing MizerParams object. Here we will illustrate the process with an 
example MizerParams object that ships with the mizer package.

I will provide a more realistic example here once it is ready
and tested.

```{r example, message=FALSE}
library(mizer)
library(mizerUR)
params <- NS_params
params@species_params$species
```

Let us introduce a detritus component. The mizerUR package provides a function
`detritus_dynamics()` that we will use to model the detritus. We set this up
by creating a named list
```{r}
dynamics <- list(detritus = "detritus_dynamics")
```

The `detritus_dynamics()` function uses two parameters 

* `detritus_external` gives the rate of change of detritus from external 
  sources in grams/year,
* `detritus_proportion` gives the proportion of the biomass consumed by fish 
  that flows into the detritus component.
  
We need to give values for these parameters in a named list
```{r}
dynamics_params <- list(`detritus_external` = 0,
                        `detritus_proportion` = 0.1)
```

Finally we need to specify the rate in proportion/year at which fish encounter 
detritus biomass. We use an allometric rate for this, which means that we only
need to specify the rate for a fish of 1 gram and rates at other sizes will
be scaled appropriately. This rate is species specific, so it goes into a new
column in the `species_params` data frame. Let us set some (unrealistic made-up)
values.
```{r}
params@species_params$rho_detritus <- 
    c(0, 0, 0, 0, 1, 5, 10, 0, 20 ,10, 12, 5) * 1e-15
```

We can now add the detritus component with `setUR()`.
```{r}
params <- setUR(params, dynamics, dynamics_params)
```
See the help page for `setUR()` for more details.

The initial value of the resource biomass is `URInitial(params)`.
We set it as follows:
```{r}
URInitial(params) <- c(detritus = 10^12)
```

We can now run the model in the usual way using the `project()` function of the
mizer package.
```{r}
sim <- project(params, progress_bar = FALSE)
```

We can extract the detritus biomass from the sim object with `UR(sim)` which
will return an array with one row for each saved time in the sim object. We
can use this in a plot for example.
```{r}
library(ggplot2)
ggplot(melt(UR(sim))) +
    geom_line(aes(x = t, y = value)) +
    scale_y_log10() +
    xlab("Time [years]") +
    ylab("Detritus biomass [g]")
```


